Computation and comparison of the stable northeastern United States marine boundary layer

### 2020-12-08 Paper/results.tex
More changes to results.tex

The comparison of the wind spectra and maximum resolvable frequency
$f_{max}$ for the different mesh resolutions is shown in figure
\ref{fig:GridStudySpectra}.  Here spectra is shown at the $z$=20 m
height and averaged using 1/3 octave-bins to enhance clarity. For the
$S_u$ and $S_w$ spectra, there are noticeable shifts in the peak
amplitude, while the peak amplitudes for $S_v$ remained fairly
constant between cases.  In all cases, the peak frequency of $S_i(f)$
also increased as the resolution increased. While the mesh resolution
was sufficient to capture the peak frequency for $S_u$, the fine
resolution with 2.5 m mesh cells was required to capture the spectral
peaks for the lateral and vertical velocities.

@@

In addition to the mean profiles described above, we also examined the
wind spectra and turbulent correlation statistics from both AMR-Wind
and Nalu-Wind for the stable 5 m/s case.  A comparison of the computed
wind spectra $S_i(f)$ from both LES codes is shown in figure
\ref{fig:CompareAMRvsNaluSpectra}.  At low frequencies, both AMR-Wind
and Nalu-Wind predict similar spectral behavior, and both codes
predict relatively similar spectral peaks for the $S_v$ and $S_w$
spectra.  For the longitudinal spectra $S_u$, however, AMR-Wind
predicts a slightly lower the peak amplitude than Nalu-Wind.

@@

The degree of stratification in each of the marine boundary layers can
be quantified through the Obukhov length $L_{Ob}$, defined here as
\begin{equation}
  L_{Ob} = -\frac{u_\tau^3 T_0}{\kappa g \langle \overline{w'T'} \rangle},
\end{equation}
where the $\kappa$=0.41 is the Kolmogorov constant, $g$ is the
gravitational acceleration, $T_0$ is the reference temperature, and
$\langle \overline{w'T'} \rangle$ is the horizontally and
time-averaged temperature flux.  In this study we adopt the same
classification guidelines as Archer et al
\cite{archer2016predominance}, and identify boundary layer cases where
$100 < L_{Ob} < 500$ as ``stable,'' and cases where $5 < L_{Ob}<100$
as ``very stable.''  From the values listed in Table
\ref{tab:CompareAMRallWS}, we find that the lower wind speeds are
``very stable,'' while the highest 15 m/s wind speed is classified as
``stable.''  This is consistent with the measured Cape Wind
distributions, where very stable conditions are more frequent than
stable conditions at lower wind speeds, and gradually disappear beyond
13 m/s (c.f. figure 5b in \cite{archer2016predominance}).

@@

  \end{tikzpicture}
  \caption{ \label{fig:SnapshotsZ20} Snapshots of the horizontal
    velocity and vertical velocity, taken at t=20,000 s of the
    simulations at height z=20 m, for the AMR-Wind calculation of the
    stable 5 m/s, 10 m/s, and 15 m/s ABL. }
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

This observation can also be seen in images taken along the flow
direction $225^\circ$, as shown in Figure
\ref{fig:SnapshotsSide}.  At the time t=20,000 s, the growth of the
of 15 m/s stable boundary layer is considerably larger than the very
stable 5 m/s and 10 m/s cases, and has nearly reached the inversion
layer height.  To reach the same level of boundary layer growth, the
very stable ABL cases would require longer simulation times, possibly
longer than the nighttime conditions measured experimentally.

@@

    \node[anchor=south west, inner sep=0](image) at (0,5.0in)
         {\includegraphics[height=2.5in]{figures/snapshots/05ms/velh_mag_side.png}};
    \node[align=left] at (2.5in, 5.0in){(a) AMR-Wind stable, 5 m/s};
  \end{tikzpicture}
  \caption{ \label{fig:SnapshotsSide} Snapshots of the horizontal
    velocity, taken at t=20,000 s along the wind direction, for
    the (a) 5 m/s, (b) 10 m/s, and (c) 15 m/s AMR-Wind stable
    calculations }

@@

Horizontally averaged flow profiles shown in Figures
\ref{fig:CompareAMRallWS} and \ref{fig:CompareAMRallTTI} provide a
more quantitative assessment of the ABL evolution.  The rapid growth
of the 15 m/s stable case leads to strong shear throughout the
boundary layer up to the inversion height, while the very stable 5 m/s
and 10 m/s cases showed minimal shear at higher elevations (figure \ref{fig:CompareAMRallWS}).  The
change in wind direction is also indicative of the boundary layer
penetration.  Approximately linear veer profiles are visible up to
heights of $z\approx$250 m and $z\approx$350 m for the 5 m/s and 10
m/s cases, respectively, while the linear veer is present in the 15
m/s stable case up to $z$=650 m.

The behavior of the TI and temperature profiles in figure
\ref{fig:CompareAMRallTTI} closely follows the shear and veer profiles
discussed above.  Despite all cases reaching similar TI levels at
$z$=20 m, the turbulence decays at different rates for each case until
reaching the top of the boundary layer.  The effect of the different
surface temperature changes is also visible in the horizontally
averaged temperature profiles.  Both the 10 m/s and 15 m/s cases impose
similar temperature changes (-1.4 K/hr and -1.5 K/hr, respectively)
and similar temperatures at the ocean surface.  However, the
cooling effect in the very stable 10 m/s case is limited to $z$<350
m, similar to the veer and turbulence profiles.

>>>>

The comparison of the wind spectra and maximum resolvable frequency
$f_{max}$ for the different mesh resolutions is shown in Figure
\ref{fig:GridStudySpectra}.  Here spectra is shown at the $z$=20 m
height and averaged using 1/3 octave-bins to enhance clarity. For the
$S_u$ and $S_w$ spectra, there are noticeable shifts in the peak
amplitude, while the peak amplitudes for $S_v$ remained fairly
constant between cases.  In all cases, the peak frequency of $S_i(f)$
also increased as the resolution increased. While the mesh resolution
was sufficient to capture the peak frequency for $S_u$, the fine
resolution with 2.5 m mesh cells was required to capture the spectral
peaks for the lateral and vertical velocities.

@@

In addition to the mean profiles described above, we also examined the
wind spectra and turbulent correlation statistics from both AMR-Wind
and Nalu-Wind for the stable 5 m/s case.  A comparison of the computed
wind spectra $S_i(f)$ from both LES codes is shown in Figure
\ref{fig:CompareAMRvsNaluSpectra}.  At low frequencies, both AMR-Wind
and Nalu-Wind predict similar spectral behavior, and both codes
predict relatively similar spectral peaks for the $S_v$ and $S_w$
spectra.  For the longitudinal spectra $S_u$, however, AMR-Wind
predicts a slightly lower the peak amplitude than Nalu-Wind.

@@

The degree of stratification in each of the marine boundary layers can
be quantified through the Obukhov length $L_{Ob}$, defined here as
\begin{equation}
  L_{Ob} = -\frac{u_\tau^3 T_0}{\kappa g \langle \overline{w'T'} \rangle},
\end{equation}
where the $\kappa$=0.41 is the Kolmogorov constant, $g$ is the
gravitational acceleration, $T_0$ is the reference temperature, and
$\langle \overline{w'T'} \rangle$ is the horizontally and
time-averaged temperature flux.  In this study we adopt the same
classification guidelines as Archer et al
\cite{archer2016predominance}, and identify boundary layer cases where
$100 < L_{Ob} < 500$ as ``stable,'' and cases where $5 < L_{Ob}<100$
as ``very stable.''  From the values listed in Table
\ref{tab:CompareAMRallWS}, we find that the lower wind speeds are
``very stable,'' while the highest 15 m/s wind speed is classified as
``stable.''  This is consistent with the measured Cape Wind
distributions, where very stable conditions are more frequent than
stable conditions at lower wind speeds, and gradually disappear beyond
13 m/s (c.f. Figure 5b in \cite{archer2016predominance}).

@@

  \end{tikzpicture}
  \caption{ \label{fig:SnapshotsZ20} Snapshots of the horizontal
    velocity and vertical velocity, taken at $t$=20,000 s of the
    simulations at height z=20 m, for the AMR-Wind calculation of the
    stable 5 m/s, 10 m/s, and 15 m/s ABL. }
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

This observation can also be seen in images taken along the flow
direction $225^\circ$, as shown in Figure
\ref{fig:SnapshotsSide}.  At the time $t$=20,000 s, the growth of the
of 15 m/s stable boundary layer is considerably larger than the very
stable 5 m/s and 10 m/s cases, and has nearly reached the inversion
layer height.  To reach the same level of boundary layer growth, the
very stable ABL cases would require longer simulation times, possibly
longer than the nighttime conditions measured experimentally.

@@

    \node[anchor=south west, inner sep=0](image) at (0,5.0in)
         {\includegraphics[height=2.5in]{figures/snapshots/05ms/velh_mag_side.png}};
    \node[align=left] at (2.5in, 5.0in){(a) AMR-Wind stable, 5 m/s};
  \end{tikzpicture}
  \caption{ \label{fig:SnapshotsSide} Snapshots of the horizontal
    velocity, taken at $t$=20,000 s along the wind direction, for
    the (a) 5 m/s, (b) 10 m/s, and (c) 15 m/s AMR-Wind stable
    calculations }

@@

Horizontally averaged flow profiles shown in Figures
\ref{fig:CompareAMRallWS} and \ref{fig:CompareAMRallTTI} provide a
more quantitative assessment of the ABL evolution.  The rapid growth
of the 15 m/s stable case leads to strong shear throughout the
boundary layer up to the inversion height, while the very stable 5 m/s
and 10 m/s cases showed minimal shear at higher elevations (Figure \ref{fig:CompareAMRallWS}).  The
change in wind direction is also indicative of the boundary layer
penetration.  Approximately linear veer profiles are visible up to
heights of $z\approx$250 m and $z\approx$350 m for the 5 m/s and 10
m/s cases, respectively, while the linear veer is present in the 15
m/s stable case up to $z$=650 m.

The behavior of the TI and temperature profiles in Figure
\ref{fig:CompareAMRallTTI} closely follows the shear and veer profiles
discussed above.  Despite all cases reaching similar TI levels at
$z$=20 m, the turbulence decays at different rates for each case until
reaching the top of the boundary layer.  The effect of the different
surface temperature changes is also visible in the horizontally
averaged temperature profiles.  Both the 10 m/s and 15 m/s cases impose
similar temperature changes (-1.4 K/hr and -1.5 K/hr, respectively)
and similar temperatures at the ocean surface.  However, the
cooling effect in the very stable 10 m/s case is limited to $z$<350
m, similar to the veer and turbulence profiles.
